### 2018-07-31 CS761-RandomizedAlgorithms/assignments/a4.tex
q5

\newpage
%Problem 5
\section{}

\newpage
%Problem 6
\section{}
\subsection{a}
\paragraph{}
For this part it will suffice to show that the Markov chain is finite, irreducible, and aperiodic. Its state space is the set of all permutations on $[n]$ so indeed it is finite. If cards $i$ and $j$ are chosen to swap, yet $i=j$ then we transition to the same permutation, and therefore the Markov chain is aperiodic. To see that it is irreducible we have to show that any permutation can be reached from any other permutation by a series of transpositions (swaps). Let $\sigma_1, \sigma_2 \in S_n$. We can write $\sigma_1 = \tau_1 \circ \dots \circ \tau_k$ and $\sigma_2 = \tau'_1 \circ \dots \circ \tau'_\ell$ where each $\tau_i$ and $\tau'_j$ is a transposition. This is a basic result in group theory. So starting from $\sigma_1$ we if we make the sequence of transitions 
$$\tau_1^{-1}, \dots, \tau_k^{-1},\tau'_\ell, \dots, \tau'_1$$
then we reach $\sigma_2$ from $\sigma_1$.
\subsection{b}
\paragraph{}
If we can demonstrate a coupling of the Markov chain with expected convergence time $O(n^2)$ then we know there exists a time $T = O(\frac{n^2}{\epsilon})$ where $\tau(\epsilon) \leq T$, at which point we can use the Coupling Lemma from Page $2$ of Lecture $18$ to achieve the desired results.
\end{document}

>>>>

	%Problem 5
	Consider a DNF formula $I$ with clauses $C_1, \dots, C_m$ and variables $x_1, \dots, x_n$. We adopt the notation from the problem statement and random process description, and further let $\text{Sat}$ denote the set of satisfying assignments for $I$. We begin by observing
	$$E[P] = E[\frac{W}{N(\sigma)}] = W\cdot E[\frac{1}{N(\sigma)}] = W\sum_{\sigma \in \text{Sat}} Pr(\sigma \text{ is chosen})\frac{1}{N(\sigma)}$$
	where the random event ``$\sigma$ is chosen" denotes the event that the random process chooses $\sigma$ to satisfy $C_i$. Since the process never chooses an assignment not in $\text{Sat}$ the sum ranges over $\sigma$ in Sat.
	We now proceed to analyze $Pr(\sigma\text{ is chosen}).$ Let $X_i$ denote the event that the random process begins by choosing clause $C_i$. We have $Pr(X_i) = \frac{w_i}{W}$ by definition. Let $Y_{i,\sigma}$ denote the event that $\sigma$ is the assignment chosen to satisfy clause $C_i$. We have
	$$Pr(Y_i\mid X_i) = \frac{p^{t_\sigma}(1-p)^{f_\sigma}}{p^{a_i}(1-p)^{b_i}} = \frac{p^{t_\sigma}(1-p)^{f_\sigma}}{w_i}$$
	where $t_\sigma$ and $f_\sigma$ denote the number of variables $\sigma$ sets to True and False respectively. Let $\text{Sat}(i)$ denote the subset of $\text{Sat}$ which satisfy $C_i$. Since
	$$\sigma\text{ is chosen} = \bigcup_i X_i \cap Y_i$$
	and the events $X_i \cap Y_i$ and $X_j \cap Y_j$ are disjoint,
	$$Pr(\sigma\text{ is chosen}) = \sum_{i \in [m]}Pr(X_i \cap Y_i) = \sum_{i:\sigma \in \text{Sat}(i)} Pr(X_i)Pr(Y_i\mid X_i) = \sum_{i: \sigma \in \text{Sat}(i)} \frac{w_i}{W} \frac{p^{t_\sigma}(1-p)^{f_\sigma}}{w_i}.$$
	Hence we have
	$$Pr(\sigma\text{ is chosen}) = \frac{N(\sigma)}{W}p^{t_\sigma}(1-p)^{f_\sigma}.$$
	Substituting this probability into our expression for $E[P]$ we see that
	$$E[P] = \sum_{\sigma \in \text{Sat}}p^{t_\sigma}(1-p)^{f_\sigma}.$$
	Indeed this is precisely the probability that the DNF formula is satisfiable.
	Let $\epsilon,\delta \in (0,1]$. We now proceed to obtain an $(\epsilon,\delta)$-approximation for the probability that the DNF formula is satisfiable. We will run the random process for $t$ trials (to be chosen later), and return the average outcome. Let $P_i = \frac{W}{N(\sigma_i)}$ be the outcome of trial $i$. Let $\mu = E[P_i]$ for all $i \in[t]$. We have previously shown that $\mu$ is precisely the probability that the DNF formula is satisfiable.
	Since $N((\sigma_i))\in [1,m]$ we have $P_i \in [\frac{W}{m}, W]$, and we see that $\frac{P_i}{W} \in [\frac{1}{m}, 1]$. Our algorithm will return
	$$\bar{P} := \frac{1}{t} \sum_{i=1}^t P_i.$$
	To be an $(\epsilon,\delta)$-approximation we want to show
	$$Pr(|\bar{P} - \mu|> \epsilon\mu) \leq \delta.$$
	We define a change of variables, so that $X_i =\frac{P_i}{W}$ and $\bar{X} = \frac{1}{t}\sum_{i=1}^t X_i$. Now showing our desired inequality for $(\epsilon,\delta)$-approximation is equivalent to showing
	$$Pr(|\bar{X} - E[\bar{X}]| > \epsilon E[\bar{X}]) \leq \delta.$$
	We now invoke the Hoeffding extension of our Chernoff Bound results in Lecture $2$, and sum both tails:
	$$Pr(|\bar{X} - E[\bar{X}]| > \epsilon E[\bar{X}]) \leq 2\exp(-2t\epsilon^2(\frac{\mu}{W})^2) \leq 2\exp(-2t(\frac{\epsilon}{m})^2) $$
	where the last inequality follows since $\frac{\mu}{W} \geq \frac{1}{m}$ (as each of $\frac{P_i}{W}$ is at least $\frac{1}{m}$).
	Hence if we choose $t$ so that
	$$t \geq \frac{-\ln(\frac{\delta}{2})}{2} \cdot (\frac{m}{\epsilon})^2$$
	then we will obtain the desired probability bound so that our algorithm is an $(\epsilon,\delta)$-approximation.$\blacksquare$
	%Problem 6
	\subsection{a}
	For this part it will suffice to show that the Markov chain is finite, irreducible, and aperiodic. Its state space is the set of all permutations on $[n]$ so indeed it is finite. If cards $i$ and $j$ are chosen to swap, yet $i=j$ then we transition to the same permutation, and therefore the Markov chain is aperiodic. To see that it is irreducible we have to show that any permutation can be reached from any other permutation by a series of transpositions (swaps). Let $\sigma_1, \sigma_2 \in S_n$. We can write $\sigma_1 = \tau_1 \circ \dots \circ \tau_k$ and $\sigma_2 = \tau'_1 \circ \dots \circ \tau'_\ell$ where each $\tau_i$ and $\tau'_j$ is a transposition. This is a basic result in group theory. So starting from $\sigma_1$ we if we make the sequence of transitions 
	$$\tau_1^{-1}, \dots, \tau_k^{-1},\tau'_\ell, \dots, \tau'_1$$
	then we reach $\sigma_2$ from $\sigma_1$.
	\subsection{b}
	If we can demonstrate a coupling of the Markov chain with expected convergence time $O(n^2)$ then we know there exists a time $T = O(\frac{n^2}{\epsilon})$ where $\tau(\epsilon) \leq T$, at which point we can use the Coupling Lemma from Page $2$ of Lecture $18$ to achieve the desired results.
	We define our joint random process for the coupling now. Let $\sigma_1$ be the permutation state of the original Markov chain, and $\sigma_2$ the permutation for the copy. Our process proceeds in stages, where stage $i$ denotes that the two permutations agree in $i$ positions. At stage $i$ we have three different types of swaps. Suppose that the original Markov chain chooses to swap cards with values $a$ and $b$. Cases:
		\item $a$ and $b$ are both in positions whose values agree in $\sigma_1$ and $\sigma_2$. Then the copy Markov chain will choose to swap $a$ and $b$ as well, preserving that we are in stage $i$.
		\item Similarly if $a$ is in a position whose values agree in $\sigma_1$ and $\sigma_2$, and $b$ is not in such a position, then the copy Markov chain will choose to swap card $b$ into $a$'s position to preserve agreement and stay in stage $i$.
		\item In final case: that $a$ and $b$ are both in non-agreement positions, we progress stages. The copy Markov chain looks at the state the original ends up in after the swap of $a$ and $b$ and chooses a non-agreement position uniformly at random, making a swap (between two non-agreement position cards) to increase the number of positions $\sigma_1$ and $\sigma_2$ agree on by at least $1$. That is, we progress at least to stage $i+1$.
	The random process ends when we are in stage $n$. This is a valid coupling since the copy Markov chain has probability of making any given swap $\frac{1}{n^2}$.
	The probability that we improve from stage $i$ to stage $\geq i+1$ during a state transition in the random process is the same as the probability that two non-agreement positions are chosen for the swap. This occurs with probability
	$$\frac{(n-i)^2}{n^2}.$$
	Hence the expected number of transitions before we move from stage $i$ to stage $i+1$ is at most
	$$\frac{n^2}{(n-i)^2}.$$
	Therefore, using linearity of expectation, the expected number of transitions before we reach stage $n$ from stage $0$ is at most
	$$\sum_{i=1}^n\frac{n^2}{(n-i)^2} \leq cn^2 = O(n^2)$$
	where $c$ is a constant, $c \approx \frac{\pi^2}{6}$. Therefore the expected convergence time of our coupling is $O(n^2)$, and we invoke the Coupling Lemma to complete the proof.$\blacksquare$
\end{document}
